First and foremost, I would like to thank my thesis advisor and committee chair, 
Christiane Jablonowski. She has provided superb mentorship and 
has contributed in a significant manner to my future career.
I would like to thank the remainder of my committee, Hans Johansen,
Mark Flanner, and Richard Rood.
I appreciate all your advice and support in this endeavor.

Further thanks to Hans Johansen for all the guidance, mentorship, 
and countless weeks spent debugging 
to get the model to work. I would also like to thank the others
at Lawerence Berkeley National Laboratory who assisted in the various facets 
of working on the model that this thesis is centered on, including
Phil Colella, Peter McCorquodale, Eli Goodfriend, and Paul Ullrich. 
Thank you for the support and helpfully answering my numerous questions.
 
 Also, thank you Cyndi McNabb, Lauren Segall, 
 Chuck Sutton, Sandra Pytlinksi, and Allison Lyons for 
 guiding me through the bureaucracy of 
 being a graduate student in two departments. Sometimes it seemed 
 like more than twice the work, so thanks for making it easy.
 
Last but far from least, I thank all my friends and family without whom 
my success and the completion of my degree would not have been possible.
Thank you to my friends and officemates in the Jablonowksi Data Lab Kevin, Paul, James, 
Weiye, Colin, Allison, and Diana for the advice and  
helpful work-related conversation. However,
I am especially grateful for the countless
hours we spent joking in the office, hanging out at conferences, and 
venturing outside the SRB for activities around Ann Arbor.
To my Applied Physics compatriots, 
thanks for the six years of adventures that only physicists can have.
Our game nights 
and random late night shenanigans made getting though Jackson a breeze.
I also thank my second family in AOSS (now CLASP) for welcoming me 
into the department and teaching me how the atmosphere and weather actually work.
You also provided much needed mental breaks from the day to day stresses of graduate school.
Of course, thank you to my long time roommate and good friend Carrie
as well as my numerous IM sports compatriots including. 
Though we established an illustrious Red Zeppelin dynasty 
in co-rec softball, we failed to win any other championships 
in the half dozen other IM sports we played.  
The fun we had playing and then recovering 
from our injuries afterwards at The Arena more than made up for those losses.
I would be remiss to forget my science policy and student government friends and mentors 
who helped turn an enjoyable side interest into a future career path.
 
 I am forever thankful for my parents who provided me with 
 countless opportunities to succeed in life as well as in academics. 
 They have always supported me in every decision I have made. 
 I would also begrudgingly thank my sister Brennan for the best 
 sibling rivalry a brother could ask for. Most importantly, I'd like to 
 thank the soon to be doctor Katie whose companionship 
 has made my time here a joy. 
 Her love and support, as well as some key proofreading,
 allowed me to get through these stressful times.